---
title: "Food Consumption Score"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = FALSE}
library(dietry)
```

The **Food Consumption Score (FCS)** is an index developed by the World Food Programme (WFP) in 1996. The **FCS** is a household level indicator that aggregates food group diversity and frequency over the past 7 days. These food groups are then weighted according to their relative nutritional value. This means that food groups that are nutritionally-dense such as animal products are given greater weight than those containing less nutritionally dense foods such as tubers. The weights are then added up to come up with a household score which are then used to classify households into either poor, borderline, or acceptable food consumption. The **FCS** is a measure of quantity of caloric intake.

## FCS questionnaire

A brief questionnaire is used to ask respondents about the frequency of their households’ consumption of eight different food groups over the previous seven days. The eight food groups are:

1. Main staples
2. Pulses
3. Vegetables
4. Fruit
5. Meat/fish
6. Milk
7. Sugar
8. Oil

Following is a model questionnaire that can be used for collecting data on **FCS**.


**Question:** I would like to ask you about all the different foods that your household members have eaten in the last 7 days. Could you please tell me how many days in the past week your household has eaten the following foods?

*(for each food, ask what the primary source of each food item eaten that week was, as well as the second main source of food, if any)*

+-----+--------------------------------------+------------+---------+----------+
|     | Food item                            | Number of  | Primary | Secondary|
|     |                                      | days eaten | source+ | source+  |
|     |                                      | in the past|         |          |
+:===:+======================================+============+=========+==========+
| 1   | Maize                                |            |         |          |
+-----+--------------------------------------+------------+---------+----------+
| 2   | Rice                                 |            |         |          |
+-----+--------------------------------------+------------+---------+----------+
| 3   | Bread/wheat                          |            |         |          |
+-----+--------------------------------------+------------+---------+----------+
| 4   | Tubers                               |            |         |          |
+-----+--------------------------------------+------------+---------+----------+
| 5   | Groundnuts and pulses                |            |         |          |
+-----+--------------------------------------+------------+---------+----------+
| 6   | Fish (eaten as a main food)          |            |         |          |
+-----+--------------------------------------+------------+---------+----------+
| 7   | Fish powder (used for flavour only)  |            |         |          |
+-----+--------------------------------------+------------+---------+----------+
| 8   | Red meat (sheep/goat/beef)           |            |         |          |  
+-----+--------------------------------------+------------+---------+----------+
| 9   | White meat (poultry)                 |            |         |          |
+-----+--------------------------------------+------------+---------+----------+
| 10  | Vegetable oil, fats                  |            |         |          |
+-----+--------------------------------------+------------+---------+----------+
| 11  | Eggs                                 |            |         |          |
+-----+--------------------------------------+------------+---------+----------+
| 12  | Milk and dairy products (main food)  |            |         |          |
+-----+--------------------------------------+------------+---------+----------+
| 13  | Milk in tea in small amounts         |            |         |          |
+-----+--------------------------------------+------------+---------+----------+
| 14  | Vegetables (including leaves)        |            |         |          |
+-----+--------------------------------------+------------+---------+----------+
| 15  | Fruits                               |            |         |          |
+-----+--------------------------------------+------------+---------+----------+
| 16  | Sweets, sugar                        |            |         |          |
+-----+--------------------------------------+------------+---------+----------+

+------------------------------------------------------------------------------+
| **+Food source codes:**                                                      |
+==============================================================================+
| Purchase = 1; Own production = 2; Traded goods/services, barter = 3;         |
| Borrowed = 4; Received as gifts = 5; Food aid = 6;                            |
| Others: (specify) = 7                                                        |
+------------------------------------------------------------------------------+

This model questionnaire should be adapted to each survey context in which it is to be used.

## Calculating the FCS

The **FCS** or the frequency weighted diet diversity score is a score calculated using the frequency of consumption of different food groups consumed by a household during the 7 days before the survey. Following are the steps to calculate the FCS:

**Step 1.** Using the standard questionnaire above, group all the food items into specific food groups (see groups in table below).

+-----+----------------------------------------+--------------------+----------+
|     | Food items                             | Food groups        | Weight   |
+:===:+========================================+====================+==========+
| 1   | Maize , maize porridge, rice, sorghum, | Main staples       | 2        | 
|     | millet pasta, bread and other cereals  |                    |          |
|     | Cassava, potatoes and sweet potatoes,  |                    |          |
|     | other tubers, plantains                |                    |          |
+-----+----------------------------------------+--------------------+----------+
| 2   | Beans, peas, groundnuts and cashew nuts| Pulses             | 3        |
+-----+----------------------------------------+--------------------+----------+
| 3   | Vegetables, leaves                     | Vegetables         | 1        |
+-----+----------------------------------------+--------------------+----------+
| 4   | Fruits                                 | Fruits             | 1        |
+-----+----------------------------------------+--------------------+----------+
| 5   | Beef, goat, poultry, pork, eggs and    | Meat and fish      | 4        |
|     | fish                                   |                    |          |
+-----+----------------------------------------+--------------------+----------+
| 6   | Milk yoghurt and other diary           | Milk               | 4        |
+-----+----------------------------------------+--------------------+----------+
| 7   | Sugar and sugar products, honey        | Sugar              | 0.5      |
+-----+----------------------------------------+--------------------+----------+
| 8   | Oils, fats and butter                  | Oil                | 0.5      |
+-----+----------------------------------------+--------------------+----------+
| 9   | Spices, tea, coffee, salt, fish powder,| Condiments         | 0        |
|     | small amounts of milk for tea.         |                    |          |
+-----+----------------------------------------+--------------------+----------+

**Step 2.** Sum all the consumption frequencies of food items of the same group, and recode the value of each group above 7 as 7.

**Step 3.** Multiply the value obtained for each food group by its weight (see food group weights in table below) and create new weighted food group scores.

**Step 4.** Sum the weighed food group scores, thus creating the food consumption score (FCS).

**Step 5.** Using the appropriate thresholds (see below), recode the variable food consumption score, from a continuous variable to a categorical variable.

+--------------------+----------------------------------------+
| **FCS**            | **Profiles**                           |
+====================+========================================+
| 0 - 21             | Poor                                   |
+--------------------+----------------------------------------+
| 21.5 - 35          | Borderline                             |
+--------------------+----------------------------------------+
| \> 35              | Acceptable                             |
+--------------------+----------------------------------------+

## Calculating the FCS using `dietry`

Steps 1, 2, 3, and 4 described above are implemented by the `dietry` package using the `fcs_recode()` and the `fcs_calculate()` functions. The `fcs_recode()` function is called within `fcs_calculate()` to perform Step 1 and 2 and then `fcs_calculate()` completes the operation by implementing steps 3 and 4. Using the `fcs01` dataset, we apply the `fcs_calculate()` function to implement these first 3 steps:

```{r fcs-calculate, eval = FALSE}
## Create a variable map to match food group labels to variables in fcs01 ----
var_map <- fcs_fg_map_variables(
  staples = "FCSStap",
  pulses = "FCSPulse",
  vegetables = "FCSVeg",
  fruits = "FCSFruit",
  meatfish = "FCSPr",
  milk = "FCSDairy",
  sugar = "FCSSugar",
  oil = "FCSFat",
  condiment = "FCSCond"
) 

## Calculate scores ----
fcs_calculate(df = fcs01, var_map = var_map, add = TRUE)
```

This provides the following output:

```{r fcs-calculate-show, echo = FALSE}
## Create a variable map to match food group labels to variables in fcs01 ----
var_map <- fcs_fg_map_variables(
  staples = "FCSStap",
  pulses = "FCSPulse",
  vegetables = "FCSVeg",
  fruits = "FCSFruit",
  meatfish = "FCSPr",
  milk = "FCSDairy",
  sugar = "FCSSugar",
  oil = "FCSFat",
  condiment = "FCSCond"
) 

## Calculate scores ----
fcs_calculate(df = fcs01, var_map = var_map, add = TRUE)
```

The output shows the original data.frame but with recoded food groups consumption frequency and with a calculated food consumption score (`fcs` variable).

Finally, to classify the food consumption scores into respective FCS classifications using the `fcs_classify()` function as follows:

```{r fcs-classify, eval = FALSE}
## Create a variable map to match food group labels to variables in fcs01 ----
var_map <- fcs_fg_map_variables(
  staples = "FCSStap",
  pulses = "FCSPulse",
  vegetables = "FCSVeg",
  fruits = "FCSFruit",
  meatfish = "FCSPr",
  milk = "FCSDairy",
  sugar = "FCSSugar",
  oil = "FCSFat",
  condiment = "FCSCond"
) 

## Calculate scores ----
fcs_df <- fcs_calculate(df = fcs01, var_map = var_map, add = TRUE)

## Classify scores ----
fcs_classify(fcs = fcs_df$fcs, add = TRUE)
```

which gives the following output:

```{r fcs-classify-show, echo = FALSE}
## Create a variable map to match food group labels to variables in fcs01 ----
var_map <- fcs_fg_map_variables(
  staples = "FCSStap",
  pulses = "FCSPulse",
  vegetables = "FCSVeg",
  fruits = "FCSFruit",
  meatfish = "FCSPr",
  milk = "FCSDairy",
  sugar = "FCSSugar",
  oil = "FCSFat",
  condiment = "FCSCond"
) 

## Calculate scores ----
fcs_df <- fcs_calculate(df = fcs01, var_map = var_map, add = TRUE)

## Classify scores ----
fcs_classify(fcs = fcs_df$fcs, add = TRUE)
```

A new data.frame is returned with the `fcs` variable and the `fcs_class` variable which indicates whether the household has a *poor*, *borderline*, or *acceptable* food consumption score.

